---
title: "Lab 1: Interaction Terms"
author: 'MTH 362: Statistical Modeling'
date: "Updated `r Sys.Date()`"
output: 
  html_document: 
    toc: TRUE
    toc_float: TRUE
    theme: cosmo
    code_download: TRUE
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

The goal of this assignment is to: (1) explore how to incorporate categorical variables into linear regression through **interaction terms**, and (2) get you acquainted with working in the RStudio/RMarkdown/`tidyverse` ecosystem.

Answer the questions below in the spaces provided (leave the "\>" sign in place). When you're finished:

1.  Change your name in the "author:" space at the top of this document.
2.  Click the "Knit" button at the top to create an HTML file with your answers.
3.  Review your answers, make any changes as necessary, and re-"Knit".
4.  Save your HTML file and upload to BlueLine.

*You can also submit a Word or PDF copy of your work if you prefer (you'll need a LaTeX compiler to create a PDF.) If you go the Word/PDF route, please pay attention to page breaks in the compiled document, you may need to mess with the formatting options.*

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

When we introduced the `penguins` data set in class, we saw that a linear regression model worked well to describe the relationship between bill length and flipper length for three species of penguins.

```{r}
library(palmerpenguins)
data(penguins)
penguins %>%
  ggplot(aes(x=bill_length_mm, 
             y=flipper_length_mm)) + 
  geom_point() + 
  geom_smooth(method='lm',
              se=FALSE) +
  labs(x='Bill length (mm)', 
       y='Flipper length (mm)', 
       title='Palmer penguins')
```

Does the "best" linear model depend on which penguin species?

```{r}
penguins %>%
  ggplot(aes(x=bill_length_mm, 
             y=flipper_length_mm)) + 
  geom_point(aes(col=species, pch=species)) + 
  geom_smooth(method='lm',
              se=FALSE) +
  labs(x='Bill length (mm)', 
       y='Flipper length (mm)', 
       title='Palmer penguins')
```

------------------------------------------------------------------------

There are two options we could pursue:

1.  Fit a linear regression model on each species separately.
2.  Use an **interaction term** to add species to a multiple linear regression model.

**Interaction**: An interaction occurs when an explanatory variable has a different effect on the outcome depending on the values of another explanatory variable.

## Question 1

Does the value of the explanatory variable `bill_length_mm` change depending on the `species`? Use a boxplot to support your answer.

> Write your response here.

```{r}
penguins %>% 
  ggplot(aes(x=species, y=bill_length_mm)) + 
  geom_boxplot(aes(fill=species)) + 
  guides(fill=FALSE) + 
  labs(x='Species', y='Petal Width')
```

You might have previously encountered interaction terms in ANOVA models, but we can use them in any linear model. The linear model with interaction looks something like this:

$$Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \beta_{12}(X_1 \times X_2) + \epsilon$$

In R, we specify an interaction term with an asterisk (`*`).

## Question 2

The code below fits the linear regression model with an additional interaction term for `species`.

a)  Based on the output, write the linear model with the necessary dummy variables.

> Write your response here.

b)  Which terms in the model are "statistically significant"?

> Write your response here.

c)  Evaluate the model assumptions.

> Write your response here.

```{r}
model_int <- lm(flipper_length_mm ~ bill_length_mm*species, data=penguins)
summary(model_int)

par(mfrow=c(2, 2))
plot(model_int)
```

Interpreting this model is a challenge! Remember, R uses reference values to code categorical variables.

## Question 3

For the Chinstrap penguin, the fitted linear model is:

$$\hat{Y}_{Chinstrap}=(158.9244 - 12.2886) + (0.7999+0.2073)X$$

a)  Create a new data set that contains only the Chinstrap penguins, and verify this result.

> Write your response here.

```{r, eval=FALSE}
# Fill in the blank, then remove the eval=FALSE option
chinstrap <- penguins %>% filter(species == _____)
model_chinstrap <- lm(flipper_length_mm ~ bill_length_mm, data=chinstrap)
summary(model_chinstrap)
```

b)  For the Adelie penguin, the fitted linear model is:

$$\hat{Y}_{Adelie} = 158.9244 + 0.799X$$

Where did these numbers come from? (Hint: Use the `summary(model_int)` output.)

> Write your response here.

c)  Create a new data set that contains only the Adelie penguins, and verify this result.

> Write your response here.

d)  Write the fitted linear model for the Gentoo penguins. Where do the numbers come from?

> Write your response here.

e)  Create a new data set that contains only the Gentoo penguins, and verify this result.

> Write your response here.

f)  What do the p-values for the interaction term represent? That is, what hypotheses are being tested?

> Write your response here.
